### 2020-08-04 manuscripts/thesis/abstract.tex
update

\begin{abstract}
Learning optimal dictionaries for sparse coding has exposed characteristic sparse features of many natural signals. However, universal guarantees of the uniqueness and stability of such features in the presence of noise are lacking. This work presents very general conditions guaranteeing when dictionaries yielding the sparsest encodings of a dataset are unique and stable with respect to measurement or modeling error. The stability constants are explicit and computable; as such, there is an effective procedure sufficient to affirm if a proposed solution to the dictionary learning problem is unique within bounds commensurate with the noise. Beyond the extension of existing results to the noisy regime, a theory of combinatorial designs for sparse supports is introduced to demonstrate that some or all generating dictionary elements are recoverable from noisy data even if the dictionary fails to satisfy the spark condition, its size is overestimated, or only a polynomial number of distinct supports appear in the encoded data. Importantly, the guarantees derived here assume no constraints on the recovered dictionary beyond a natural upper bound on its size. The work closes with a lengthy discussion of the results and some remaining open challenges seeding future research directions.
\end{abstract}