### 2020-08-13 manuscripts/thesis/thesis.tex
redid abstract

%% thesis.tex 2014/04/11
%
% Based on sample files of unknown authorship.
%
% The Current Maintainer of this work is Paul Vojta.
%
% To compile this file, run "latex thesis", then "biber thesis"
% (or "bibtex thesis", if the output from latex asks for that instead),
% and then "latex thesis" (without the quotes in each case).

\documentclass{ucbthesis}

% *** CITATION STUFF ***
%\usepackage{biblatex} 
%\bibliography{references} 
% the above doesn't work, using this instead:
\usepackage[backend=bibtex,style=numeric]{biblatex}
\addbibresource{references}

% *** MATH STUFF ***
\usepackage{amsmath, amssymb, amsthm} 
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{conjecture}{Conjecture}
\newtheorem{problem}{Problem}
\newtheorem{question}{Question}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}

% *** ALIGNMENT STUFF***
\usepackage{array}
\usepackage{bm}
\usepackage{rotating} % provides sidewaystable and sidewaysfigure
\usepackage{url}

% Double spacing, if you want it.  Do not use for the final copy.
% \def\dsp{\def\baselinestretch{2.0}\large\normalsize}
% \dsp

% If the Grad. Division insists that the first paragraph of a section
% be indented (like the others), then include this line:
% \usepackage{indentfirst}

\addtolength{\abovecaptionskip}{\baselineskip}

%\hyphenation{mar-gin-al-ia}
%\hyphenation{bra-va-do}

\begin{document}

% Declarations for Front Matter

%\title{On the uniqueness and stability of dictionaries for sparse representation of noisy signals}
\title{A Tale of Two Dictionary Learning Problems}
\author{Charles Garfinkle}
\degreesemester{Summer}
\degreeyear{2020}
\degree{Doctor of Philosophy}
\chair{Friedrich Sommer}
\othermembers{Bruno Olshausen \\
  Mike Deweese \\
  Bin Yu}
% For a co-chair who is subordinate to the \chair listed above
% \cochair{Professor Benedict Francis Pope}
% For two co-chairs of equal standing (do not use \chair with this one)
% \cochairs{Professor Richard Francis Sony}{Professor Benedict Francis Pope}
\numberofmembers{4}
% Previous degrees are no longer to be listed on the title page.
% \prevdegrees{B.A. (University of Northern South Dakota at Hoople) 1978 \\
%   M.S. (Ed's School of Quantum Mechanics and Muffler Repair) 1989}
\field{Neuroscience}
% Designated Emphasis -- this is optional, and rare
% \emphasis{Colloidal Telemetry}
% This is optional, and rare
% \jointinstitution{University of Western Maryland}
% This is optional (default is Berkeley)
% \campus{Berkeley}

% For a masters thesis, replace the above \documentclass line with
% \documentclass[masters]{ucbthesis}
% This affects the title and approval pages, which by default calls this
% document a "dissertation", not a "thesis".

\maketitle
% Delete (or comment out) the \approvalpage line for the final version.
\approvalpage
\copyrightpage

\include{abstract}

\begin{frontmatter}

\begin{dedication}
\null\vfil
%\begin{center} 
I want to dedicate this to my dad. At the same time, I don't want to, because the results of this work can't be matched to the time with him I traded to produce them, and it feels as though dedicating this work to him would in some sense be measuring his life up against them. 
%These theorems, while fundamental, are just something people might cite to feel better about the sparse coding analyses they were already going to do, anyway. And yet, despite total freedom offered by the Helen Wills Neuroscience Institute, I worried the brain was so dauntingly complex that I should feel lucky to have at last been offered a well-defined relevant problem to work on, particularly one I felt even remotely equipped to solve myself, so as to feel deserving of a Ph.D. from UC Berkeley. 
The fact is that whatever may be the eventual butterfly effect of this work on the scientific ecosystem, these mathematical truths could have waited, arrived at their own pace, via myself or another, and I could have been more present with my dad as he waited patiently for a terminal cancer to take his life.

While I couldn't really fathom what it actually meant for him to die -- to never see or talk to or hug him again -- the idea of putting original work out there, of crafting my own piece of the overwhelming puzzle that is neuroscience, was alluring. Yet, I feel as though it was a feeling of inadequacy, of having to prove myself worthy of the position I had come to find myself in after a first failed project that largely drove these efforts.

Let's just say, then, that this dedication is the first bit of a life's work, real work devoid of any prestigious position and the associated self-doubt, pressure, or delusions of grandeur; a life dedicated to passionate ideas, to solving problems I care about and that really matter, to loving and caring for others or, as my dad would say, to ``being a mensch". May this dedication serve to hold me publicly accountable to that. Also I love you, mom. I dedicate and don't dedicate this and everything to come to you, too.

\vspace{12pt}
%\end{center}
\vfil\null
\end{dedication}

% You can delete the \clearpage lines if you don't want these to start on
% separate pages.

\tableofcontents
\clearpage
\listoffigures
\clearpage
%\listoftables

\begin{acknowledgements}
First and foremost, I would like to thank Chris Hillar for clearing the path in his original work on the dictionary problem, and for vetting all of my proofs many, many times over. He asked for nothing in return for his time, attention, and mathematical expertise, and continues to extend this generosity to anyone nerdy enough to handle it. Fritz Sommer for recommending my acceptance to Berkeley and for advising me as his student; although I eventually had to leave his lab to work on other problems of interest to me, he kindly lent me his time over the course of four years. The other members of my thesis committee: Bruno Olshausen, Mike DeWeese, and Bin Yu, for their time and consideration over the years as well, and for approving this dissertation. My mom would have killed me if I didn't claim this degree, so they saved my life. The graduate program managers of the Helen Wills Neuroscience Institute: Tony Leonard and Candace Groskreutz for their kindness and patience, and to the HWNI itself. The Redwood Center for Theoretical Neuroscience and the friends I made in the lab; I hope we remain in each other's lives. Finally, a heartfelt thank you to Eric Weiss for pointing me to the student co-ops, and the Berkeley Student Cooperative (www.bsc.coop) for the work they do providing the amazing experience of living co-operatively in Berkeley. The co-ops changed my life, and I am forever grateful -- thank you so, so much to the hundreds of talented, creative, weird, passionate, intelligent, kind, thoughtful, absurd, inspiring people I have lived with throughout the years. CZ or die. 
\end{acknowledgements}

\end{frontmatter}

\pagestyle{headings}

%\section{Foreword}
%This thesis presents results published Dec.~2019 in a paper of the same name in \emph{IEEE Transactions on Signal Processing}, largely verbatim. Readers should refer to the Discussion and Future Directions as well as Figures \ref{probpigeon} and \ref{reasonableC2} for additional insights not offered in the paper.

% (Optional)
 %\part{Main Text}

\include{introduction}
\include{results}
\include{proofs}
\include{discussion}
%\include{appendix}

\printbibliography

\end{document}