### 2021-10-19 manuscripts/L1_old/L1.tex

\documentclass[journal, onecolumn]{IEEEtran}

% *** MATH PACKAGES ***
\usepackage{amsmath, amssymb, amsthm} 
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{conjecture}{Conjecture}
\newtheorem{problem}{Problem}
\newtheorem{question}{Question}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}

\usepackage[pdftex]{graphicx}

% *** ALIGNMENT PACKAGES ***
\usepackage{array}
\usepackage{cite}
\usepackage{bm}

% correct bad hyphenation here
%\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

\title{$\ell_1$-norm minimization}
\author{Charles~J.~Garfinkle}

\maketitle

\begin{problem}\label{OptimizationProblem}
Find matrices $\mathbf{B}$ and vectors \mbox{$\mathbf{\overline x}^{(1)}, \ldots, \mathbf{\overline x}^{(N)}$} solving:
\begin{align}\label{minsum}
\min \sum_{i = 1}^N \|\mathbf{\overline x}^{(i)}\|_0 \ \
\text{subject to} \ \ \|\mathbf{z}^{(i)} - \mathbf{B}\mathbf{\overline x}^{(i)}\|_2 \leq \eta_0, \ \text{for all $i$}
\end{align}
by solving:
\begin{align}\label{minsum}
\min \sum_{i = 1}^N \|\mathbf{\overline x}^{(i)}\|_1 \ \
\text{subject to} \ \ \|\mathbf{z}^{(i)} - \mathbf{B}\mathbf{\overline x}^{(i)}\|_2 \leq \eta_1, \ \text{for all $i$}.
\end{align}
\end{problem}

\begin{proof}[Proof for $k=1$]
Since the only 1-uniform hypergraph with the SIP is $[m]$, which is obviously regular, we require only $\mathbf{x}^{(i)} = c_i \mathbf{e}_i$ for $i \in [m]$, with $c_i \neq 0$ to guarantee  linear independence. While we have yet to define $C_1$ generally, in this case we may set $C_1 = 1/ \min_{\ell \in [m]} |c_{\ell}|$. % L_2(\mathbf{A})  \min_{\ell \in [m]} |c_{\ell}| - \max_{\ell \in [m]} |c_\ell|$.

Fix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfying $L_2(\mathbf{A}) > 0$, since here we have $2\mathcal{H} = {[m] \choose 2}$, and suppose some matrix $\mathbf{B}$ and vectors $\mathbf{\overline x}^{(i)} \in \mathbb{R}^{\overline m}$ have  $\|\mathbf{A}\mathbf{x}^{(i)} - \mathbf{B}\mathbf{\overline x}^{(i)}\|_2 \leq \varepsilon$ for all $i$. Note that $\mathbf{\overline x}^{(i)} \neq 0$, since otherwise we would reach the following contradiction: $\|\mathbf{A}_i \|_2 \leq C_1 |c_i| \|\mathbf{A}_i \|_2  \leq \|\mathbf{Ax}_i\|_2 \leq C_1\varepsilon \leq  (?) < \min_{\ell \in [m]} \|\mathbf{A}_\ell\|_2$. %< L_2(\mathbf{A}) \leq L_1(\mathbf{A}) = \min_{i \in [m]} \|\mathbf{A}_{i}\|_2$.

Let $\mathbf{\overline x}^{(i)} = \sum_{j=1}^m \overline c^{(i)}_j \mathbf{e}_j$ and let $\pi: [m] \to [\overline m]$ be the map $\pi(i) = \arg \max_j \overline c^{(i)}_j$. By the triangle inequality,
\begin{align}\label{1D}
\|c_i\mathbf{A}_i - \overline{c}^{(i)}_{\pi(i)}\mathbf{B}_{\pi(i)}\|_2  -  \|\sum_{k \neq \pi(i)} \mathbf{B}_k \overline c^{(i)}_k \|_2
\leq \|\mathbf{A}\mathbf{x}^{(i)} - \mathbf{B}\mathbf{\overline x}^{(i)}\|_2 
\leq \varepsilon
\end{align}

Hence, %assuming $\|\mathbf{B}\|_2 \leq \|\mathbf{A}\|_2$ for all $i \in [m]$,
\begin{align}\label{1D}
%(|c_i| - |\overline{c}^{(i)}_{\pi(i)}|) \|\mathbf{A}_i\|_2 
%= \|\mathbf{A}_i (c_i - \overline{c}^{(i)}_{\pi(i)})\|_2 
%\leq 
\|c_i\mathbf{A}_i - \overline{c}^{(i)}_{\pi(i)}\mathbf{B}_{\pi(i)}\|_2 
&\leq \varepsilon + \|\sum_{k \neq \pi(i)} \mathbf{B}_k \overline c^{(i)}_k \|_2 \\
%&= \varepsilon + \|\mathbf{B}_{\pi(j)} \overline c^{(i)}_{\pi(j)} + \sum_{k \neq \pi(i) \neq \pi(j)} \mathbf{B}_k \overline c^{(i)}_k\|_2 \\
%&\leq \varepsilon + \|\mathbf{B}_{\pi(j)} \overline c^{(i)}_{\pi(j)} - c_j\mathbf{A}_j\|_2 + \|c_j\mathbf{A}_j - \sum_{k \neq \pi(i) \neq \pi(j)} \mathbf{B}_k \overline c^{(i)}_k\|_2 \\
&\leq \varepsilon + \|\mathbf{B}\|_2 \| \mathbf{\overline x}^{(i)}_{\pi(i)=0} \|_1 \\
&= \varepsilon + \|\mathbf{A}\|_2 \| \mathbf{\overline x}^{(i)}_{\pi(i)=0} \|_1 \\
%&\leq \varepsilon + \|\mathbf{A}\|_2 (|c_i| - |\overline c_{\pi(i)}^{(i)}|)
%\leq \varepsilon +  \|\mathbf{\overline x}^{(i)} \|_1
%\leq \varepsilon + |c_i|
\end{align}

%Split x's in two, scale them, use A and B?

%Or for $m=2$, 
%\begin{align}\label{1D}
%\|c_1\mathbf{A}_1 - \overline{c}^{(1)}_{\pi(1)}\mathbf{B}_{\pi(1)}\|_2 
%&\leq \varepsilon + \| \mathbf{B}_k \overline c^{(1)}_{k \neq \pi(1)} \|_2 \\
%\|c_2\mathbf{A}_2 - \overline{c}^{(2)}_{\pi(2)}\mathbf{B}_{\pi(2)}\|_2 
%&\leq \varepsilon + \| \mathbf{B}_k \overline c^{(2)}_{k \neq \pi(2)} \|_2 
%\end{align}

%We now show that the second term is also controlled by $\varepsilon$ (due to the fact that the other columns of the dictionary must also be coded for?).

%\begin{align}\label{1D}
%|\overline{c}^{(j)}_\ell| \|c_{i}\mathbf{A}_{i} - \overline{c}^{(i)}_\ell \mathbf{B}_{\ell}\|_2 
%-
%|\overline{c}^{(i)}_\ell| \|c_{j}\mathbf{A}_{j} - \overline{c}^{(j)}_\ell\mathbf{B}_{\ell}\|_2 
%&\leq \varepsilon + \|\mathbf{A}\|_2 \| \mathbf{\overline x}^{(i)}_{\pi(i)=0} \|_1 \\
%&\leq \varepsilon + \|\mathbf{A}\|_2 (|c_i| - |\overline c_{\pi(i)}^{(i)}|)
%\leq \varepsilon +  \|\mathbf{\overline x}^{(i)} \|_1
%\leq \varepsilon + |c_i|
%\end{align}

%\begin{align}\label{1D}
%|\overline{c}^{(i)}_\ell| \|c_{j}\mathbf{A}_{j} - \overline{c}^{(j)}_\ell\mathbf{B}_{\ell}\|_2  + |\overline{c}^{(j)}_\ell| \|c_{i}\mathbf{A}_{i} - \overline{c}^{(i)}_\ell \mathbf{B}_{\ell}\|_2 
%&\leq (|\overline{c}^{(i)}_\ell|  +  |\overline{c}^{(j)}_\ell|) \varepsilon 
%+ |\overline{c}^{(j)}_\ell| \|\sum_{k \neq \ell} \mathbf{B}_k \overline c^{(i)}_k \|_2
%+ |\overline{c}^{(i)}_\ell| \|\sum_{k \neq \ell} \mathbf{B}_k \overline c^{(j)}_k \|_2 \\
%&\leq \varepsilon + \|\mathbf{A}\|_2 \| \mathbf{\overline x}^{(i)}_{\pi(i)=0} \|_1 \\
%&\leq \varepsilon + \|\mathbf{A}\|_2 (|c_i| - |\overline c_{\pi(i)}^{(i)}|)
%\leq \varepsilon +  \|\mathbf{\overline x}^{(i)} \|_1
%\leq \varepsilon + |c_i|
%\end{align}

We now show that $\pi$ is injective (in particular, a permutation if $\overline m = m$). Suppose that $\pi(i) = \pi(j) = \ell$ for some $i \neq j$ and $\ell$. Then we have: %Scaling and summing these inequalities by $|\overline{c}_{i}|$ and $|\overline{c}_{j}|$, respectively, and applying the triangle inequality to annihilate the terms in $\mathbf{B}_\ell$, we obtain:
\begin{align*}
(|\overline{c}^{(i)}_\ell| + |\overline{c}^{(j)}_\ell|) \left(\varepsilon + \|\mathbf{A}\|_2 \| \mathbf{\overline x}^{(i)}_{\pi(i)=0} \|_1 \right) 
%+ \|\mathbf{A}\|_2 \left( |\overline{c}^{(j)}_\ell| (|c_i| - |\overline{c}^{(i)}_\ell| ) + |\overline{c}^{(i)}_\ell| (|c_j| - |\overline{c}^{(j)}_\ell| ) \right)
&\geq |\overline{c}^{(i)}_\ell| \|c_{j}\mathbf{A}_{j} - \overline{c}^{(j)}_\ell\mathbf{B}_{\ell}\|_2  + |\overline{c}^{(j)}_\ell| \|c_{i}\mathbf{A}_{i} - \overline{c}^{(i)}_\ell \mathbf{B}_{\ell}\|_2 \nonumber \\
&\geq \|\mathbf{A}(\overline{c}^{(i)}_\ell c_{j} \mathbf{e}_{j} - \overline{c}^{(j)}_\ell c_{i}\mathbf{e}_{i})\|_2 \nonumber \\ 
&\geq \sqrt{2}  L_2(\mathbf{A}) \|\overline{c}^{(i)}_\ell c_{j} \mathbf{e}_{j} - \overline{c}^{(j)}_\ell c_{i}\mathbf{e}_{i}\|_2 \nonumber \\
&\geq  L_2(\mathbf{A}) (|\overline{c}^{(i)}_\ell| + |\overline{c}^{(j)}_\ell|) \min_{\ell \in [m]} |c_\ell |,
\end{align*}
hence, assuming $\|\mathbf{A}\|_2 = 1$,
\begin{align*}
\varepsilon  &\geq  L_2(\mathbf{A})  \min_{\ell \in [m]} |c_\ell | - \| \mathbf{\overline x}^{(i)}_{\pi(i)=0} \|_1 \\
&=
\end{align*}
contradicting our assumed upper bound on $\varepsilon$. Hence, the map $\pi$ is injective and so $\overline m \geq m$. %Setting $\overline J = \pi([m])$ and

Letting $\mathbf{P}$ and $\mathbf{D}$ be the $\overline m \times \overline m$ permutation and invertible diagonal matrices with, respectively, columns $\mathbf{e}_{\pi(i)}$ and $\frac{\overline{c}_i}{c_i}\mathbf{e}_i$ for $i \in [m]$ (otherwise, $\mathbf{e}_{i}$ for $i \in [\overline{m}] \setminus [m]$), we may rewrite \eqref{1D} to see that for all $i \in [m]$:
\begin{align*}
\|\mathbf{A}_i - (\mathbf{BPD})_i\|_2 
= \|\mathbf{A}_i - \frac{\overline{c}_i}{c_i}\mathbf{B}_{\pi(i)}\|_2 
\leq \frac{\varepsilon + |c_i|}{|c_i|} 
\leq C_1\varepsilon + 1.
\end{align*}
%\vspace{-.2 cm}
\end{proof}

\end{document}